---
title: "County analysis"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{county-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Let's look at how many pills have gone to each county in West Virginia.

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# Uncomment and run the lines below to see if you have the packages required already installed
# packages <- c("dplyr", "ggplot2", "forcats", "jsonlite", "knitr", "geofacet", "scales")
# if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
#   install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org") # }

library(arcos)
library(knitr)
library(tigris)
library(viridis)
library(dplyr)
library(ggplot2)
library(scales)
library(forcats)
```

```
# Uncomment and run the lines below to see if you have the packages required already installed
# packages <- c("tidyverse", "jsonlite", "knitr", "geofacet", "scales", "forcats")
# if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
#   install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org") # }


# These are all the packages you'll need to run everything below 

library(arcos)
library(knitr)
library(tigris)
library(viridis)
library(dplyr)
library(ggplot2)
library(scales)
library(forcats)
```

First, let's look at the pharmacies in Mingo, West Virginia.

The total pills per pharmacy in a county can be pulled in with `total_pharmacies_county()`.

```{r top_county_pharmacy}
mingo <- total_pharmacies_county(county = "Mingo", state="WV", key="WaPo")

kable(head(mingo))
```


Which were the most prolific pharmacies in Mingo between 2006 and 2014?

Let's chart them out.

```{r top_county_pharmacy_chart, fig.width=9, fig.height=7}
ggplot(mingo,
       aes(x=total_dosage_unit, y=fct_reorder(buyer_name, total_dosage_unit))) +
  geom_segment(
       aes(x = 0,
           y=fct_reorder(buyer_name, total_dosage_unit),
           xend = total_dosage_unit,
           yend = fct_reorder(buyer_name, total_dosage_unit)),
           color = "gray50") +
           geom_point() +
    scale_x_continuous(label=comma) +
  labs(x="Dosage units", y="", 
       title = "Pills sold at Mingo, WV pharmacies",
       subtitle = "Between 2006 and 2014",
       caption = "Source: The Washington Post, ARCOS") +
  theme_minimal()
```

Okay, now we can look at all the counties in West Virginia.

Pull that data with `summarized_county_annual()`.

```{r wv_data}
wv <- summarized_county_annual(state="WV", key="WaPo")

kable(head(wv))
```

For easy mapping, we'll use Census shape files pulled with the [**Tigris** package](https://github.com/walkerke/tigris).

```{r tigris, warning=F, message=F, quietly=T, results="hide"}
## Set the option for shapefiles to load with sf
options(tigris_class = "sf")

## Function to download county shapefiles in West Virginia
wv_shape <- counties(state="WV", cb=T)

## Join the county dosage data we pulled
wv <- left_join(wv, wv_shape, by=c("countyfips"="GEOID"))
```

```{r facet_map, fig.width=9, fig.height=7}
# Mapping with ggplot2, sf, and viridis

wv %>%
  ggplot(aes(geometry=geometry, fill = DOSAGE_UNIT, color = DOSAGE_UNIT)) +
  facet_wrap(~year, ncol=2) +
  geom_sf() +
  coord_sf(crs = 26915) + 
  scale_fill_viridis(direction=-1, label = comma) +
  scale_color_viridis(direction=-1, label = comma) +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) +
  labs(title="Oxycodone and hydrocodone pills in West Virginia", caption="Source: The Washington Post, ARCOS")
```

Looks nice. You should probably [adjust for population](https://wpinvestigative.github.io/arcos/articles/annual-maps.html) next.

```{r pop, warning=F, message=F, fig.width=9, fig.height=7}

population <- county_population(state="WV", key="WaPo") %>% 
# isolate the columns so it doesn't conflict in a join (there are doubles, that's why)
    select(countyfips, year, population)

left_join(wv, population) %>% 
  mutate(per_person=DOSAGE_UNIT/population) %>%
  ggplot(aes(geometry=geometry, fill = per_person, color = per_person)) +
  facet_wrap(~year, ncol=2) +
  geom_sf() +
  coord_sf(crs = 26915) + 
  scale_fill_viridis(direction=-1, label = comma) +
  scale_color_viridis(direction=-1, label = comma) +
  theme_void() +
  theme(panel.grid.major = element_line(colour = 'transparent')) +
  labs(title="Oxycodone and hydrocodone pills in West Virginia per person", caption="Source: The Washington Post, ARCOS")

