---
title: "Per capita pharmacies"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{per-capita-pharmacies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Here's our process for calculating pills per person within a county— figures that end up in many of our [articles](https://www.washingtonpost.com/investigations/the-opioid-crisis-15-percent-of-the-pharmacies-handled-nearly-half-of-the-pills/2019/08/12/b24bd4ee-b3c7-11e9-8f6c-7828e68cb15f_story.html).

First, let's load up the packages we need.

There's a lot of new ones listed here so we can make fancy tables.

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Uncomment and run the lines below to see if you have the packages required already installed
# packages <- c("dplyr", "ggplot2", "jsonlite", "knitr", "geofacet", "scales")
# if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
#   install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org") # }

library(arcos)
library(dplyr)
library(ggplot2)
library(lubridate)
library(data.table)
library(formattable)
library(stringr)
library(scales)
library(knitr)
```

```
# Uncomment and run the lines below to see if you have the packages required already installed
# packages <- c("dplyr", "jsonlite", "knitr", "geofacet", "scales")
# if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
#   install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org") # }

library(arcos)
library(dplyr)
library(lubridate)
library(data.table)
library(formattable)
library(vroom)
library(stringr)
library(scales)
library(knitr)
```

Let's look at all pharmacies in West Virginia and their total oxycodone and hydrocodone pill orders.

And use the `total_pharmacies_state()` function.

```{r pull_data}
west_virginia <- total_pharmacies_state(state="WV", key="WaPo")

kable(head(west_virginia))
```

Looks nice. We've got total_dosage_unit and total_records. We can ignore total_records for this example.

Each pharmacy is listed, as well as which county it resides in.

To help normalize the number of pills ordered, we need to get each county's population in West Virginia.

We can use the `county_population()` function.

```{r population}
population <- county_population(state="WV", key="WaPo")

kable(head(population))
```


Alright, we have population tables but they're annual figures.

We have one number for dosage per pharmacy. We don't need seven years of population data.

So we'll average it out.

And then join it to the pharmacy list.

```{r population_refined}
population <- population %>% 
  group_by(BUYER_COUNTY, BUYER_STATE, countyfips) %>% 
  # Figure out the average population between available years
  summarize(average_population=mean(population, na.rm=T)) %>% 
  ## Have to quickly rename these columns to make them lower case so they'll join easily to the other data frame
  rename(buyer_county=BUYER_COUNTY, buyer_state=BUYER_STATE)

## Join the data
wv_joined <- left_join(west_virginia, population)

kable(head(wv_joined))
```

Now that we've joined the county population data to each pharmacy, we can do some math to figure out the pills per person per year.

That's pills / average population / 7

```{r join}
wv_joined <- wv_joined %>% 
  mutate(per_person=total_dosage_unit/average_population/7)

kable(head(wv_joined))
```

Let's make sure we're only dealing with chain and retail pharmacies.

This requires some supplemental data from `buyer_addresses()` and `not_pharmacies()`

```{r colors}

## Get a list of addresses because it includes BUYER_BUS_ACT information
pharmacy_list <- buyer_addresses(state="WV", key="WaPo")

# We just want the BUYER_BUS_ACT to tell if these are practitioners are retail pharmacies
# This will help us filter out the appropriate pharmacies

pharmacy_list <- pharmacy_list %>% 
  select(buyer_dea_no=BUYER_DEA_NO, BUYER_BUS_ACT)

# Join to the original data set
wv_joined <- left_join(wv_joined, pharmacy_list)

# Filter the data so we only have retail and chain pharmacies
wv_joined <- wv_joined %>% 
  filter(BUYER_BUS_ACT=="RETAIL PHARMACY" | BUYER_BUS_ACT=="CHAIN PHARMACY")

# Just in case, let's get the BUYER_DEA_NO of pharmacies that aren't really pharmacies
not_pharms <- not_pharmacies(key="WaPo") %>% pull(BUYER_DEA_NO)

# Filter those out, too, if they're in there
wv_joined <- wv_joined %>% 
  filter(!buyer_dea_no %in% not_pharms)

# clean up column names so we can make a pretty table
wv_joined <- wv_joined %>% 
  select(Pharmacy=buyer_name, City=buyer_city, County=buyer_county, `County population`=average_population,
         Pills=total_dosage_unit, `Pills per person`=per_person) %>% 
  mutate(`County population`=round(`County population`),
         `Pills per person`=round(`Pills per person`, 1)) %>% 
  arrange(desc(`Pills per person`)) %>% 
  slice(1:100)

# Create some custome colors
customGreen0 = "#DeF7E9"
customGreen = "#71CA97"
customRed = "#ff7f7f"

# produce a table
wv_joined %>% 
  formattable(align=c("l", "l", "l", "r", "r", "r"),
              list(Pharmacy = formatter("span", style = ~ style(color="grey", font.weight = "bold")),
                   Pills=color_tile(customGreen0, customGreen),
                   `Pills per person` = normalize_bar(customRed)
                   ))
```


