---
title: "Suspicious orders detection"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{suspicious-order-detection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The DEA is purposefully vague in how it defines what is a suspicious amount of opioids ordered by a buyer.

But in a [recent lawsuit](https://www.docketbird.com/court-documents/In-re-National-Prescription-Opiate-Litigation/Exhibit-12-Rafalski-Report-Redacted/ohnd-1:2017-md-02804-01876-014), a former DEA agent described five methodologies:

* Maximum  Monthly, Trailing 6 Month Threshold
* 2x Trailing 12 Month average
* Extraordinary Order Method – 3x Trailing 12 Month Average
* Maximum 8,000 Dosage Units Monthly
* Maximum  Daily  Dosage  Units

This vignette will show how to analyze one pharamacy's ordering patterns to identify suspicious orders using the first two listed methodologies.

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Uncomment and run the lines below to see if you have the packages required already installed
# packages <- c("dplyr", "ggplot2", "jsonlite", "knitr", "geofacet", "scales", "zoo", "knitr")
# if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
#   install.packages(setdiff(packages, rownames(installed.packages())), repos = "http://cran.us.r-project.org") # }

library(arcos)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(stringr)
library(scales)
library(zoo)
library(knitr)
```

We're going to look at a Walgreens pharmacy in Oviedo that was part of a [half dozen pharmacies](https://www.orlandosentinel.com/news/os-xpm-2012-04-06os-walgreens-dea-oviedo-20120406-story.html) in Florida sanctioned by the DEA in 2011 for suspicious activity. The Post later determined at the height of the opioid crisis, [Walgreens handled nearly one in five of the most addictive opioids](https://www.washingtonpost.com/investigations/2019/11/07/height-crisis-walgreens-handled-nearly-one-five-most-addictive-opioids/?arc404=true).

## Investigate 

First, we need to find the BUYER_DEA_NO for that store. All we know of [this specific pharmacy](https://www.wftv.com/news/local/dea-inspects-oviedo-walgreens-oxycodone-probe/287945741/) is the city and the address: Oviedo and Lockwood Boulevard. 

Let's load up the data on buyers. The city of Oviedo is in Seminole County, Florida. Let's search based on that.

```{r buyer, eval=F}
pharm <- buyer_addresses(county = "Seminole", state="FL", key="WaPo")

glimpse(pharm)
```

```{r buyer_real, echo=F}
pharm <- readRDS("data/seminole_pharms.RDS")

glimpse(pharm)
```

Okay, we have 468 pharmacies in Seminole County. Are there any on Lockwood?

```{r lockwood}
pharm <- filter(pharm, 
                grepl("LOCKWOOD", BUYER_ADDRESS1))

kable(pharm)
```

Okay, we have four. And there are two that specifically Walgreens. 

It's either BW8487438	or FW4552091. They both have the same address.

`pharmacy_raw(buyer_dea_no = "FW4552091", key = "WaPo")` returns an error from the ARCOS api, so `BW8487438` must be the correct BUYER_DEA__NO.


```{r oviedo, eval=F}
pharm1 <- pharmacy_raw(buyer_dea_no = "BW8487438", key = "WaPo")

glimpse(pharm1)
```

```{r oviedo_real, echo=F}
pharm1 <- readRDS("data/solo_pharm.RDS")

glimpse(pharm1)
```

These are the raw results from this one pharmacy. More than 4,000 orders.

Let's narrow this list down to only Oxycodone and by month.

```{r summarizing}
# consolidating by month

pharm_monthly <- pharm1 %>% 
  # setting transactio code to S because those are purchases
  filter(TRANSACTION_CODE=="S") %>% 
  filter(DRUG_NAME=="OXYCODONE") %>% 
  mutate(TRANSACTION_DATE=mdy(TRANSACTION_DATE)) %>% 
  mutate(year=year(TRANSACTION_DATE),
         month=month(TRANSACTION_DATE),
         month_date=mdy(paste0(month, "/1/", year))) %>% 
  group_by(year, month, month_date) %>% 
  summarize(pills=sum(DOSAGE_UNIT))
```

## Maximum Monthly, Trailing 6 Month Threshold

We use the `lag()` function to pull six months of pill orders prior and then keep the largest order from that list.

```{r six_month_max, warning=F, message=F}
# Slices out the max orders within a window of 6 months prior to the current month
monthly_drug_six <- pharm_monthly %>% 
  ungroup() %>% 
  arrange(month_date) %>% 
  mutate(lag1=lag(pills, 1),
         lag2=lag(pills, 2),
         lag3=lag(pills, 3),
         lag4=lag(pills, 4),
         lag5=lag(pills, 5),
         lag6=lag(pills, 6)
  ) %>% 
  pivot_longer(cols=5:10, names_to="lag", values_to="total") %>% 
  group_by(year, month, pills, month_date) %>% 
  arrange(desc(total)) %>% 
  slice(1)  %>% 
  ungroup()


# creating a column that indicates if a monthly dosage exceeds the 6 month max prior

monthly_drug_six <- monthly_drug_six %>% 
  mutate(six_max_flag=case_when(
         pills > total ~ T,
         TRUE ~ F))

#monthly_drug_six$trail_six_max_flag <- ifelse(monthly_drug_six$pills > monthly_drug_six$total, 1, 0)
```

```{r six_month_plot, warning=F, message=F}

ggplot(monthly_drug_six, aes(x=month_date, y=pills)) +
  geom_col(fill="cadetblue3") +
  geom_step(aes(x=month_date, y=total, fill="Maximum 6 month order"), color="tomato", opacity=.6)+
    scale_y_continuous(label=comma) +
  theme_minimal() +
  labs(title="Monthly oxycodone pill orders from a Walgreens pharmacy in Oviedo, FL",
       subtitle='As compared to the DEA suspicious "exceeds the 6-month maximum" order criteria')

```

How often did this store's orders exceed the 6-month max order?

```{r table1}
table(monthly_drug_six$six_max_flag)
```

## 2x Trailing 12 Month average

Next, we'll determing what double the trailing 12-month average is.

```{r rolling avg, warning=F, message=F}
# we'll use the rollmean() function from the zoo package

rolling <- pharm_monthly %>% 
  ungroup() %>% 
  arrange(month_date) %>% 
  mutate(avg_pills = zoo::rollmean(pills, k = 12, fill = NA, align="right")) %>% 
  mutate(avg_pills_2x=avg_pills*2) %>% 
  mutate(roll_flag_2x=case_when(
    avg_pills_2x<pills ~ T,
    TRUE ~ F
  ))

# CALCULATE ROLLING AVERAGE HERE

ggplot(rolling, aes(x=month_date, y=pills)) +
  geom_col(fill="cadetblue3") +
  geom_step(aes(x=month_date, y=avg_pills_2x, fill="Double the 12 month rolling average"), direction="mid",color="tomato", opacity=.06) +
  scale_y_continuous(label=comma) +
  theme_minimal() +
  labs(title="Monthly oxycodone pill orders from a Walgreens pharmacy in Oviedo, FL",
       subtitle='As compared to the DEA suspicious "double the 12-month rolling average" order criteria')

```



How often did this store's orders exceed double the 12-month rolling average?

```{r table2}
table(rolling$roll_flag_2x)
```

